---
title: "FG_size"
format: html
editor: visual
---

## Functional Group by Size Model

```{r}
library(ppjsdm)
library(dplyr)
library(tidyr)
library(ggpubr)
library(patchwork)
library(tidyr)
```

\
Load data

```{r}
data <- read.csv("../../../data/ausplots/data_cleaned.csv")
```

Load species functional groups (FG)

```{r}
fg <- read.csv("../../../data/ausplots/species_class.csv")
```

If only doing visualisation can load the df that this first part makes:

```{r}
fg_size_df <- read.csv("fg_size_df.csv")
```

#### Select threshold for size

To set up a function group by size model we first need make groups into their function group (FG), and then use this to divide individuals into relatively large or relatively small individuals of the FG.

First attach the species_classes FG to data_cleaned

```{r}
#something strange happened to df so get rid of weird encoding 
fg <- fg %>% 
  mutate(Genus_Species = iconv(Genus_Species, to = "ASCII", sub = " "))

#join into df 
data1 <- left_join(data, fg, 
                   by = "Genus_Species")

data1 %>% count(Class) #some species are not in the fg csv so nas
```

Check what species have an NA for class

```{r}
nas <- data1 %>% filter(is.na(Class)) 

nas %>% count(Genus_Species, sort = TRUE)
```

Most are looking like understorey trees but we can plot the diameter of them quickly:

```{r}
nas %>% filter(! Genus_Species == "Unidentified tree") %>% 
  ggplot() + 
  geom_histogram(aes(x = Diameter)) + 
  theme_bw()
```

\
Can compare this to the range of diameters (sizes) in species we have given an FG to already:

```{r}
d2 <- data1 %>% filter(!is.na(Class))  

diam <- gghistogram(d2,
            x = "Diameter", 
            y = "count", 
            fill = "Class",
            bins = 50, 
            rug = TRUE,
            add_density = FALSE) +
        xlim(0, 200)

diam
```

Right, as expected flat distribution of canopy and more left skewed (one bin has most individuals) of subcanopy. A quick look at the species that haven't been given an FG shows that the ones with the biggest diameters are still understorey trees. So, we can hardcode all of the leftover species into the subcanopy group. We keep the unidentified trees as unidentified as we can use this as a misc class:

```{r}
data1 <- data1 %>% 
  mutate(Class = if_else(Genus_Species == "Unidentified tree", "Unid tree",Class)) %>% 
  mutate(Class = if_else(is.na(Class), "Subcanopy", Class))

data1 %>% count(Class)
```

Check the distribution of subcanopy to canopy individuals:

```{r}
data1 %>% count(Class)
```

There are three sites with enough unidentified trees to be important:

- Bruxner 147 (out of 853)

- Osullivans 128 (out of 514)

- Tinebank 123 (out of 454)

We will leave these as is for the moment.

### Run PPJSDM for all sites

So, the function to output the fit with a group being a fg x site:

```{r}
fg_size_sites <- function(site, #single site 
                       show_size_freq = TRUE,
                       config_only = FALSE, #if TRUE, returns config only and exits function
                       threshold = 15, 
                       short_range = 10, 
                       short_model = "exponential"){ #threshold to exclude a group 

  df <- data1 %>%
    filter(Site_Name == site)
  
  
  #jitter coordinates
  df <- df %>%
    mutate(
      is_duplicated = n() > 1, #create column of TRUE/FALSE 
      #new_column_name = if_else(condition, true, false): so condition=column name, if true=fill with, if false=fill with
      x_jitter = if_else(is_duplicated, Ausplot_X + runif(n(), -0.025, 0.025), Ausplot_X), #create x_jitter column
      y_jitter = if_else(is_duplicated, Ausplot_Y + runif(n(), -0.025, 0.025), Ausplot_Y) #create y_jitter column
    ) 
  
  
  #make groups of fg x size  
    d <- df %>%
      group_by(Class) %>% 
      mutate(median_diameter = ceiling(median(Diameter, na.rm = TRUE)) + 3.5)  %>% 
      mutate(size = case_when(
                                Diameter < median_diameter ~ "small", 
                                Diameter >= median_diameter ~ "large")) %>% 
      ungroup() %>% 
      mutate(new_group = paste0(Class, sep = " ", size)) %>% 
      group_by(new_group) %>% 
      mutate(observation_count = n()) %>% 
      ungroup() 
    d %>% count(new_group)
     
      filter(!(observation_count < threshold))
 
    
  #make config
  configuration <- ppjsdm::Configuration(d$x_jitter,
                                         d$y_jitter, 
                                         types = d$new_group)
  plot(configuration)
  
  if(config_only){
    return(configuration)
  }
  
  d %>% count(new_group)
  
   #set parameters 
  window <- ppjsdm::Rectangle_window(x_range = c(0, 100), 
                                     y_range = c(0, 100))
  
  nspecies <- length(levels(configuration$types))
  
set.seed(1030)  
  #fit model 
  fit<- ppjsdm::gibbsm(configuration = configuration, #do the fit 
                       window = window,
                       short_range = matrix(10, nspecies, nspecies), 
                       model = "exponential",
                       saturation = 10, 
                       nthreads = 4, 
                       use_regularization = FALSE, 
                       fitting_package = "glmnet",
                       dummy_distribution = "stratified",
                       min_dummy = 1, dummy_factor = 1e10, 
                       max_dummy = 1e3)

  
  sum <- summary(fit)
  sum
  return(list(fit = fit, sum = sum, config = configuration))
}
```

In FG + Size Model, these are the sites that are impacted by changing the threshold. Of these, even fewer are changed by moving the threshold between 10 and 15.

ANU363: 1 subcanopy large, 5 subcanopy small \
Candelo: 4 subcanopy small \
Carey: 1 subcanopy large\
Collins: 12 subcanopy large \
Frankland: 9 subcanopy large\
Goodenia: 11 subcanopy large \
Herberton: 4 subcanopy large\
Newline: 2 subcanopy small \
Sutton: 3 subcanopy large \
WaratahMix: no subcanopy anyways \
Warren: 3 subcanopy large\
WogWay: 3 subcanopy large, 12 subcanopy small

Before running function need to source the make_summary_fun

```{r}
source("../../../functions/make_summary_fun.R")
```

Run function for all sites in ausplots:

```{r}
sites <- unique(data1$Site_Name)

full_df <- data.frame(matrix(ncol = 9, nrow = 0))
colnames(full_df) <- c("from", "to", "alpha", "lo", "hi", "lo_numerical",
                       "hi_numerical", "Potential","site")

set.seed(3)
for (i in sites){
  
  fg_size_mod <- fg_size_sites(site = i, #single site 
                         show_size_freq = FALSE,
                         config_only = FALSE, #if TRUE, returns config only and exits function
                         threshold = 15, 
                         short_range = 10, 
                         short_model = "exponential")
  
  working_df <- make_sum_df(fits = list(fg_size_mod$fit), #use make_summary_df function to output summary as df 
                            summ = list(fg_size_mod$sum))
  
  working_df <- working_df %>% mutate(site = i)                              
  
  full_df<- rbind(full_df, working_df)
}


```

First, need to check if any species that didn't have a class assigned snuck through:

```{r}
full_df %>% count(from)

full_df %>% count(to)

full_df %>% count(site)
```

Check if there are unreasonable CIs and there's none.

```{r}
full_df<- full_df %>% 
  mutate(range_ci = hi - lo)

a <- full_df %>% filter(range_ci > 5)
a %>% count(site)
```

#### Prepare dataframe for visualisation

Make some new columns to make visualisation a bit easier:

```{r}
fg_size_df <- full_df %>% 
  filter(! from %in% c("Unid tree small", "Unid tree large")) %>% 
  mutate(size_from = gsub("^[^ ]*", "", from)) %>% 
  mutate(size_to = gsub("^[^ ]*", "", to)) %>% 
  mutate(class_from = gsub(" .*", "", from)) %>% 
  mutate(class_to = gsub(" .*", "", to)) %>% 
  mutate(int = paste0(from, "_", to)) %>% 
  mutate(size_int = paste0(size_from, "", size_to))
```

Add new column to delimitate for within a functional group and between a functional group:

```{r}
fg_size_df <- fg_size_df %>% 
  mutate(which = if_else(int %in% c("Subcanopy small_Subcanopy small", "Subcanopy small_Subcanopy large", "Subcanopy large_Subcanopy large", "Canopy small_Canopy small", "Canopy small_Canopy large", "Canopy large_Canopy large"), 
                         "within", 
                         "between"))

fg_size_df %>% count(which)
```

Check size ints

```{r}
fg_size_df %>% count(size_int)

fg_size_df <- fg_size_df %>% 
  mutate(size_int = case_when(size_int %in% 
                                c(" small large"," large small", "small large") ~ "small large", 
                              size_int == " large large" ~ "large large", 
                              size_int == " small small" ~ "small small", 
                              TRUE ~ size_int)) 
                              
```

Save out df

```{r}
write.csv(fg_size_df, "fg_size_df_t15_updated.csv")
```

### Visualisation

Let's spilt so within v between a functional group for better understanding (and comparison to the species model):

```{r}
#| fig-width: 8
#| fig-height: 8


fg_size_df %>% filter(!from %in% c("Unid tree small", "Unid tree large")) %>% 
  filter(! to %in% c("Unid tree small", "Unid tree large")) %>% 
ggplot(
       aes(x = alpha, 
           y = int)) + 
  geom_boxplot() + 
  geom_point(alpha = 0.25) + 
  theme_bw() + 
  geom_vline(xintercept = 0, colour = "red") + 
  facet_wrap(~which)
```

Can also make an overall boxplot similar to current Fig 1 of intraspecific interactions with FG put in later.

Let's look at intra-group first:

```{r}
#| fig-width: 9

fg_size_mod_within <- fg_size_df %>% 
  filter(!from %in% c("Unid tree small", "Unid tree large")) %>% 
  mutate(size_int = case_when(
    size_int == " small small" ~ "Small ↔ Small", 
    size_int == "small large" ~ "Small ↔ Large", 
    size_int == " large large" ~ "Large ↔ Large")) %>% 
  filter(which == "within") %>% 
  mutate(int = as.factor(int)) %>% 
  mutate(int = fct_relevel(int, "Subcanopy large_Subcanopy large", 
                           "Subcanopy small_Subcanopy large", 
                            "Subcanopy small_Subcanopy small",
                           "Canopy small_Canopy small", 
                           "Canopy small_Canopy large", 
                           "Canopy large_Canopy large"
                           )) %>% 
  mutate(size_int = as.factor(size_int)) %>% 
  mutate(size_int = fct_relevel(size_int, "Small ↔ Small", 
                                 "Small ↔ Large", 
                                "Large ↔ Large")) %>% 
  mutate(class_from = as.factor(class_from)) %>% 
  mutate(class_from = fct_relevel(class_from, "Subcanopy", "Canopy")) %>% 
  ggplot(aes(x = alpha, y = size_int, groups = class_from)) + 
  geom_violin(colour = "black", 
              scale = "width", 
              width = 0.7,
              fill = NA,
              position = position_dodge(0.75)) +
  geom_boxplot(colour = "black", position = position_dodge(0.75),
               outlier.shape = NA, 
               fill = NA,
               width = 0.4) + 
  geom_point(aes(colour = class_from), 
             position = position_jitterdodge(seed = 21, 
                                             jitter.width = 0.3,
                                             dodge.width = 0.75), 
             shape = 19, 
             size = 1.75, 
             alpha = 0.5) +
  geom_vline(xintercept = 0, colour = "red") +
   # stat_summary(aes(fill = size_int), fun = mean,
   #             geom = "point",
   #             shape = 19,
   #             size = 3, 
   #             color="black",
   #             position = position_dodge(width=0.7, preserve = "single"), 
   #             show.legend = FALSE) + 
  scale_colour_manual(values = c("#39568CFF", "#35B779FF"), 
                      labels = c("Canopy", "Subcanopy"),
                      limits = c("Canopy", "Subcanopy"),
                      name = "Strata") +
  ylab("") +
  theme_bw() + 
  ggtitle("FG Size Model") +
  xlim(c(-1.5, 1.5))

fg_size_mod_within

ggsave("fg_size_mod_within.png", width = 10)
```

This is really, really similar to the species_size model.

```{r}
fg_size_df %>% 
  filter(!from %in% c("Unid tree small", "Unid tree large")) %>% 
  count(site)
```

We can make a similar one for between fgs:

```{r}
fg_size_df %>% 
  filter(!from %in% c("Unid tree small", "Unid tree large")) %>% 
  mutate(size_int = case_when(
    size_int == " small small" ~ "Small ↔ Small", 
    size_int == "small large" ~ "Small ↔ Large", 
    size_int == " large large" ~ "Large ↔ Large")) %>% 
  filter(which == "between") %>% 
  mutate(int = as.factor(int)) %>% 
  ggplot(aes(x = alpha, y = rev(int))) + 
  geom_violin(colour = "black", 
              scale = "width", 
              width = 0.7,
              fill = NA,
              position = position_dodge(0.75)) +
  geom_boxplot(colour = "black", position = position_dodge(0.75),
               outlier.shape = NA, 
               fill = NA,
               width = 0.4) + 
  geom_point(position = position_jitterdodge(seed = 21, 
                                             jitter.width = 0.3,
                                             dodge.width = 0.75), 
             shape = 19, 
             size = 1.75, 
             alpha = 0.5) +
  geom_vline(xintercept = 0, colour = "red") +
   stat_summary(aes(colour = int), fun = mean,
               geom = "point",
               shape = 19,
               size = 3, 
               color="blue",
               show.legend = FALSE) + 
  theme_bw() + 
  ylab("") +  xlim(c(-1, 1))
```

This model reduces the number of interaction groups possible as there is no inter v intraspecific interaction layer on top of the within v between functional group. Therefore there is only 4 between functional group interactions. All of the interesting interactions seen in the species_size model have disappeared - espeically the one between subcanopy large - canopy small , subcanopy large - canopy large and canopy small - subcanopy large. This would change our findings quite a bit. However, this is only if we are using traditional weighted means, any other metric (mean or inverse of se weighted mean) and the mean comes in to the median and there is not much difference between the two models.

We can also understand at the site level what is happening. First add the region into the df:

```{r}
fg_size_df <- fg_size_df %>% 
   mutate(georegion = case_when(site %in% c("Weeaproinah", "Turtons", "Lardner") ~ "S_VIC", 
                               site %in% c("ANU101", "ANU363", "ANU589", "Ada Tree", "HardyCreek") ~ "N_VIC", 
                               site %in% c("Dawson", "Frankland", "Clare", "Giants") ~ "S_WA",
                               site %in% c("Carey", "Dombakup", "Warren",  "Sutton","Collins") ~ "N_WA", 
                               site %in% c("Baldy", "Koombooloomba", "Lamb Range", "Herberton") ~ "QLD", 
                               site %in% c("MinesRd", "A-Tree", "BirdTree", "BlackBull", "Lorne", "Tinebank", "Bruxner", "Osullivans") ~ "N_NSW", 
                               site %in% c("Newline", "WaratahMix", "WogWay", "Goodenia", "Candelo") ~ "S_NSW", 
                               site %in% c("BenRidge", "Caveside", "Mackenzie", "MtField", "MtMaurice", "NorthStyx") ~ "d_TAS", 
                               site %in% c("BondTier", "BlackRiver", "Weld", "MtField", "ZigZag", "Supersite", "Bird", "Flowerdale", "Dip") ~ "o_TAS")) %>% 
   mutate(region = case_when(georegion %in% c("S_VIC", "N_VIC", "S_NSW") ~ "SE_AUS", 
                            georegion %in% c("S_WA", "N_WA") ~ "WA",
                            georegion %in% c("QLD", "N_NSW") ~ "N_AUS",  
                            georegion %in% c("o_TAS", "d_TAS") ~ "SE_AUS"))
```

```{r}
#| fig-height: 12

fg_size_df %>%
  filter(!from %in% c("Unid tree small", "Unid tree large")) %>% 
  filter(class_from == class_to) %>% 
  ggplot(aes(x = alpha, y = site, colour = class_from)) + 
  facet_grid(region~size_int, scales = "free") +
  geom_point(size = 2.5) + 
  scale_colour_manual(values = c("darkgreen", "lightgreen"), name = "") +
  geom_vline(xintercept = 0, colour = "red") + 
  theme_bw(base_size = 13)
```

### Comparison

So we have a functional group - size model. We need to compare this to a model where we fit species size and then add the functional group on later.

We can do a simple quantitative comparison. We can make some summary tables for the functional_group_size model first:

```{r}
fg_size <- fg_size_df %>% 
  filter(!from %in% c("Unid tree small", "Unid tree large")) %>% 
  group_by(int) %>% 
  summarise(fg_mean = mean(alpha), 
            fg_median = median(alpha), 
            fg_min = min(alpha), 
            fg_max = max(alpha)) %>% 
  ungroup
```

Do the same for the species_size model. However this will end up being a combination of inter + intraspecific interactions therefore it makes sense that is is so similar to the fg_size model. Therefore we can spilt this one up further:

```{r}
df <- read.csv("../species_diameter/df_add3.5_t15.csv")

sp_size <- df %>% 
  group_by(group) %>% 
    summarise(sp_mean = mean(alpha), 
            sp_median = median(alpha), 
            sp_min = min(alpha), 
            sp_max = max(alpha)) %>% 
  ungroup

sp_size_intra <- df %>% 
  filter(from == to) %>% 
  group_by(group) %>% 
    summarise(sp_mean = mean(alpha), 
            sp_median = median(alpha), 
            sp_min = min(alpha), 
            sp_max = max(alpha)) %>% 
  ungroup


sp_size_inter <- df %>% 
  filter(!from == to) %>% 
  group_by(group) %>% 
    summarise(sp_mean = mean(alpha), 
            sp_median = median(alpha), 
            sp_min = min(alpha), 
            sp_max = max(alpha)) %>% 
  ungroup

```

Easiest way to compare is to merge the two dfs. To do that need to fix the group names and rename cols:

```{r}
fg_size <- fg_size %>% 
  mutate(int = gsub(" ", ".", int))

fg_size <- fg_size %>% 
  rename(group = int)

match(fg_size$group, sp_size$group)

#rename some levels 
fg_size <- fg_size %>% 
  mutate(group = case_when(group == "Canopy.small_Canopy.large" ~
                       "Canopy.large_Canopy.small", 
                       group == "Subcanopy.large_Canopy.large" ~ "Canopy.large_Subcanopy.large", 
                       group == "Subcanopy.large_Canopy.small" ~ "Canopy.small_Subcanopy.large", 
                       group == "Subcanopy.small_Canopy.large" ~ "Canopy.large_Subcanopy.small", 
                       group == "Subcanopy.small_Canopy.small" ~ "Canopy.small_Subcanopy.small", 
                       group == "Subcanopy.small_Subcanopy.large" ~ "Subcanopy.large_Subcanopy.small", 
                       TRUE ~ group))
```

Merge the dfs.

```{r}
compares <- merge(fg_size, sp_size, by = "group") #overall very similar

compares <- merge(fg_size, sp_size_intra, by = "group") #intraspecific very similar to fg 

compares <- merge(fg_size, sp_size_inter, by = "group") #interspecific starts to be more different from fg 
```

```{r}
## grab df for quick visualisation 
fg_size_df <- read_csv("scripts/model_specifications/fg_size/fg_size_df.csv")

#see what groups exist 
groups <- fg_size_df %>% count(int)

#get rid of unid trees lol 
fg_size_df <- fg_size_df %>% 
  filter(! class_from == "Unid") %>% 
  filter(! class_from == "Unid")

#rename for plotting
fg_size_df <- fg_size_df %>% 
  mutate(size_int = case_when(size_int == "small small" ~ "Small ↔ Small", 
                               size_int == "small large" ~ "Small ↔ Large", 
                               size_int == "large large" ~ "Large ↔ Large")) %>% 
  mutate(size_int = as.factor(size_int)) %>% 
  mutate(size_int = fct_relevel(size_int, "Small ↔ Small", 
                                 "Small ↔ Large", 
                                 "Large ↔ Large"
  )) %>% 
  mutate(int = as.factor(int))

ggplot(fg_size_df, 
       aes(x = alpha, 
           y = int)) + 
  geom_boxplot() + 
  geom_point(alpha = 0.06) + 
  facet_grid(rows = vars(which), 
             scales = "free") + 
  geom_vline(xintercept = 0, colour = "black", linetype = "dashed") + 
  theme_bw() + 
  ylab("")

fg_size_df %>%
  mutate(sig = ifelse(lo > 0 | hi < 0, 1 , NA)) %>% 
  count(sig)

f <- fg_size_df %>% 
  group_by(int) %>% 
  summarise(median = median(alpha))

```
