---
title: "covars_correlation"
format: html
editor: visual
---

## Correlation of Model Specifications to Covariates

### Covariate Data

```{r}
#Load site covariates 
site_covars <- read.csv("../data/site_covariates.csv")

#Add regions 
site_bioclim1 <- site_bioclim1 %>% 
mutate(region = Site_Name) %>% 
  mutate(region = if_else(region %in% c("Carey", "Dombakup", "Warren", "Dawson", "Giants", "Sutton", "Frankland", "Clare", "Collins"), 
                          "WA", region)) %>% 
  mutate(region = if_else(region %in% c("ANU101", "ANU363", "ANU589", "Ada Tree", "HardyCreek", "Weeaproinah", "Turtons", "Lardner"), 
                          "SE_VIC", region)) %>% 
  mutate(region = if_else(region %in% c("Newline", "WaratahMix", "WogWay", "Goodenia", "Candelo"), 
                          "SE_NSW", region)) %>% 
  mutate(region = if_else(region %in% c("MinesRd", "A-Tree", "BirdTree", "BlackBull", "Lorne", "Tinebank", "Bruxner", "Osullivans"), 
                          "N_NSW", region)) %>% 
  mutate(region = if_else(region %in% c("Baldy", "Koombooloomba", "Lamb Range", "Herberton"), 
                          "QLD", region)) %>% 
  mutate(region = if_else(region %in% c("BenRidge", "Caveside", "Mackenzie", "MtField", "MtMaurice", "NorthStyx"), 
                          "d_TAS", region)) %>% 
  mutate(region = if_else(region %in% c("BondTier", "BlackRiver", "Weld", "Weld", "MtField", "ZigZag", "Supersite", "Bird", "Dip", "Flowerdale"), "o_TAS", region)) 

```

We should look at correlation before looking at correlation

```{r}
pairs(~MAT + MAP + FPI + GPP + yearlymaxNDVI + ETa + AI, 
      data = site_covars)

ggplot(data = site_covars, 
       aes(x = FPI, 
            y = GPP, 
            colour = region)) + 
  geom_point()
```

### Correlation of alpha coefs to covariates

Let's get the c-c and s-s coefs from the Crown Class model specification

```{r}
df_cc <- final_ccmod %>% 
  filter(species_from == "Eucalyptus") %>% 
  filter(species_to == "Eucalyptus") %>% 
  filter(class_from == class_to) %>% 
  filter(class_from == c("Suppressed")) %>% 
  filter(class_to == c("Suppressed"))

df_cc <- df_cc %>% rename(Site_Name = site)

site_covars1 <- merge(x = site_covars1, y = df_cc[, c("Site_Name", "alpha")], by = "Site_Name", all.x = TRUE)

site_covars1 <- site_covars1 %>% rename(s = alpha)
```

```{r}
df <- site_covars1 %>%
  pivot_longer(
    cols = -c(Site_Name, cd, s, Longitude, Latitude, region),           # everything except the ID column
    names_to = "covariate",  # new column with variable names
    values_to = "value"     # new column with values
  )

df1 <- df %>%
  filter(covariate %in% c("MAP", "ETa"))

df2 <- df %>%
  filter(covariate %in% c("MAT","FPI","GPP" , "yearlymNDVI", "stand_basal_area", "ETa"))


ggplot(data = df2, 
       aes(x = value, 
           y = cd)) + 
  geom_smooth(method = "lm") + 
  geom_point(size = 2) +
  facet_wrap(~covariate, scales = "free_x") +
 # xlab("Within C/D Coef (CC)") + 
  ylab("") + 
  theme_bw()

ggplot(data = df2, 
       aes(x = value, 
           y = cd, 
           colour = region)) + 
  #geom_smooth(method = "lm") + 
  geom_point(size = 2) +
  facet_wrap(~covariate, scales = "free_x") +
 # xlab("Within C/D Coef (CC)") + 
  ylab("") + 
  theme_bw()
```

```{r}
d <- df2 %>% filter(covariate == "FPI")
d <- na.omit(d)

f <- lm(cd ~ value, data = d)
summary(f)
```

```{r}
ggplot(data = df2, 
       aes(x = s, 
           y = value,
           colour = region)) + 
  #geom_hline(yintercept = 0, colour = "red", linewidth = .75, linetype = 3) + 
  geom_point(size = 2) +
  facet_wrap(~covariate, scales = "free_y") +
  xlab("Within S Coef (CC)") + 
  ylab("") + 
  theme_bw()
```

Let's do the same thing for the SpCC model

```{r}
df_spcc <- final_spcc_mod %>% 
  filter(!species_from == "Non-euc") %>% 
  filter(!species_to == "Non-euc") %>% 
  filter(class_from == class_to) %>% 
  filter(class_from == c("Co/dominant")) %>% 
  filter(class_to == c("Co/dominant"))

df_spcc <- df_spcc %>% rename(Site_Name = site)
df_spcc <- df_spcc %>% dplyr::select(Site_Name, alpha)

cov_spcc <- site_covars %>%
  left_join(df_spcc, by = "Site_Name")

cov_spcc_cd <- cov_spcc %>% rename(cd = alpha)

################### S
df_spcc <- final_spcc_mod %>% 
  filter(!species_from == "Non-euc") %>% 
  filter(!species_to == "Non-euc") %>% 
  filter(class_from == class_to) %>% 
  filter(class_from == c("Suppressed")) %>% 
  filter(class_to == c("Suppressed"))

df_spcc <- df_spcc %>% rename(Site_Name = site)
df_spcc <- df_spcc %>% dplyr::select(Site_Name, alpha)

cov_spcc_s <- site_covars %>%
  left_join(df_spcc, by = "Site_Name")

cov_spcc_s <- cov_spcc_s %>% rename(s = alpha)
```

```{r}
cov_spcc_cd1 <- cov_spcc_cd %>%
  pivot_longer(
    cols = -c(Site_Name, cd, Longitude, Latitude, region),           # everything except the ID column
    names_to = "covariate",  # new column with variable names
    values_to = "value"     # new column with values
  )

df1 <- df %>%
  filter(covariate %in% c("MAP", "ETa"))

cov_spcc_cd1 <- cov_spcc_cd1 %>%
  filter(covariate %in% c("MAT","FPI","GPP" , "yearlymaxNDVI", "yearlymeanNDVI", "stand_basal_area", "ETa"))


ggplot(data = cov_spcc_cd1, 
       aes(x = cd, 
           y = value,
           colour = region)) + 
  #geom_hline(yintercept = 0, colour = "red", linewidth = .75, linetype = 3) + 
  geom_point(size = 2) +
  facet_wrap(~covariate, scales = "free_y") +
  xlab("Within C/D Coef (SpCC)") + 
  ylab("") + 
  theme_bw()+ 
  xlim(c(-2.5, 0.5))
```

```{r}
cov_spcc_s1 <- cov_spcc_s %>%
  pivot_longer(
    cols = -c(Site_Name, s, Longitude, Latitude, region),           # everything except the ID column
    names_to = "covariate",  # new column with variable names
    values_to = "value"     # new column with values
  )


cov_spcc_s1 <- cov_spcc_s1 %>%
  filter(covariate %in% c("MAT","FPI","GPP" , "yearlymaxNDVI", "stand_basal_area", "ETa"))


ggplot(data = cov_spcc_s1, 
       aes(x = s, 
           y = value,
           colour = region)) + 
  #geom_hline(yintercept = 0, colour = "red", linewidth = .75, linetype = 3) + 
  geom_point(size = 2) +
  facet_wrap(~covariate, scales = "free_y") +
  xlab("Within S Coef (SpCC)") + 
  ylab("") + 
  theme_bw()+ 
  xlim(c(-2.5, 0.5))
```

### PCA

Attempting PCA of CC model c/d and s with various covariates. First get the df

```{r}
df_cc <- final_ccmod %>% 
  filter(species_from == "Eucalyptus") %>% 
  filter(species_to == "Eucalyptus") %>% 
  filter(class_from == class_to) %>% 
  filter(class_from == c("Co/dominant")) %>% 
  filter(class_to == c("Co/dominant"))

df_cc <- df_cc %>% rename(Site_Name = site)

site_covars1 <- merge(x = site_covars1, y = df_cc[, c("Site_Name", "alpha")], by = "Site_Name", all.x = TRUE)

site_covars1 <- site_covars1 %>% rename(cd = alpha)
```

Select some covariates - for C/D

```{r}
site_covars2 <- site_covars1 %>% dplyr::select(MAT,FPI, yearlymaxNDVI, stand_basal_area, ETa, cd)
site_covars2 <- na.omit(site_covars2)
```

Load libraries

```{r}
library(vegan)
library(stats)
library(factoextra)

pca <- prcomp(site_covars2, scale = TRUE, center = TRUE)
fviz_pca_var(pca, repel = TRUE)

pca$rotation
fviz_contrib(pca, choice = "var", axes = 2)
```

For S

```{r}
site_covars2 <- site_covars1 %>% dplyr::select(MAT,FPI,GPP , yearlymaxNDVI, stand_basal_area, ETa, s)
site_covars2 <- na.omit(site_covars2)

pca <- prcomp(site_covars2, scale = TRUE, center = TRUE)
fviz_pca_var(pca, repel = TRUE)

pca$rotation
fviz_contrib(pca, choice = "var", axes = 1)
```

Something interesting may be going on in the FPI and yearlymaxNDVI. Let's check if out before we write it all off.

There is a significant negative correlation between repulsion and FPI IF we get rid of the N NSW highly productive but low repulsion sites

```{r}
fpi <- df2 %>% filter(covariate == "FPI")
fpi <- fpi %>% filter(value < 19)

ggplot(data = fpi, 
       aes(x = value, 
           y = cd)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm") +
 # xlab("Within C/D Coef (CC)") + 
  ylab("") + 
  theme_bw() + 
  xlim(c(7, 19))

f <- lm(cd ~ value, data = fpi)
summary(f)
```

Without the two outliers more than 0.95, there is a significant correlation

```{r}
ndvi <- df2 %>% filter(covariate == "yearlymaxNDVI")
ndvi <- ndvi %>% filter(value <0.95)

ggplot(data = ndvi, 
       aes(x = value, 
           y = cd)) +
  geom_point(size = 2) +
geom_smooth(method = "lm") +
 # xlab("Within C/D Coef (CC)") + 
  ylab("") + 
  theme_bw() 

f <- lm(cd ~ value, data = ndvi)
summary(f)
```

### Correlation of alpha coefs to bioclim

Get the bioclims we care about

```{r}
site_bioclim1 <- site_bioclim %>% 
  dplyr::select(Site_Name, AnnualMeanTemp, TempSeasonality, AnnualPrecipitation,
                PrecipSeasonality, PrecipColdestQuarter)

```

Let's use the CC Model (seems to be better)

```{r}
df_cc <- final_ccmod %>% 
  filter(species_from == "Eucalyptus") %>% 
  filter(species_to == "Eucalyptus") %>% 
  filter(class_from == class_to) %>% 
  filter(class_from == c("Suppressed")) %>% 
  filter(class_to == c("Suppressed"))

df_cc <- df_cc %>% rename(Site_Name = site)

site_bioclim1 <- merge(x = site_bioclim1, y = df_cc[, c("Site_Name", "alpha")], by = "Site_Name", all.x = TRUE)

site_bioclim1 <-site_bioclim1 %>% rename(cd = alpha.x)
```

```{r}
df_bio <- site_bioclim1 %>%
  pivot_longer(
    cols = -c(Site_Name, cd, s, region),           # everything except the ID column
    names_to = "covariate",  # new column with variable names
    values_to = "value"     # new column with values
  )

ggplot(data = df_bio, 
       aes(x = cd, 
           y = value,
  #geom_hline(yintercept = 0, colour = "red", linewidth = .75, linetype = 3) + 
  geom_point(size = 2) +
  facet_wrap(~covariate, scales = "free_y") +
  xlab("Within C/D Coef (CC)") + 
  ylab("") + 
  theme_bw()
```

```{r}
ggplot(data = df_bio, 
       aes(x = s, 
           y = value,
           colour = region)) + 
  #geom_hline(yintercept = 0, colour = "red", linewidth = .75, linetype = 3) + 
  geom_point(size = 2) +
  facet_wrap(~covariate, scales = "free_y") +
  xlab("Within C/D Coef (CC)") + 
  ylab("") + 
  theme_bw()
```
