---
title: "covars_correlation"
format: html
editor: visual
---

## Correlation of Alpha (Species + Size) to Covariates

### Load Data

```{r}
#Load site covariates 
site_covars <- read.csv("site_covariates_new.csv")

#Load alpha data 
sp_size <- read.csv("../model_specifications/species_diameter/df_add3_5_t15.csv")
```

\
So our working hypothesis is that variables that relate to the productivity of a site would influence how much competition occurs. Competition is strongest in the within-species (within-group) variables.

Within-species there are two FGs (subcanopy and canopy) and three size interacting groups (L-L, S-L, S-S). Therefore there is 6 possible groups that we might think are affected differently by the site variables.

However, let's start with our hypothesis: the strongest competition is in canopy L-L group and we think that the strength of this competition is affected by site variables. We can run a correlation test on this:

```{r}
can_l <- sp_size |> 
  filter(species_from == species_to) |> 
  filter(class_from == 'Canopy') |> 
  filter(size_int == 'large_large') |> 
  group_by(site) |> 
  summarise(mean_alpha = mean(alpha), 
            median_alpha = median(alpha), 
            max_alpha = max(alpha)) |> 
  ungroup()

#Clare doesnt have a canopy l-l group so need to remove this group from the variables
vars2 <- vars |> filter(! Site_Name == "Clare")

cor_l <- cor(x = can_l$mean_alpha, 
             y = vars2[3:29]) 

corrplot(cor_l)
```
