---
title: "make_site_covariates"
format: html
editor: visual
---

```{r}
library(terra)
library(geodata)
library(dplyr)
library(MODISTools)
library(vegan)
library(tidyr)
```

## Make Site Covariates

For the 48 AusPlot sites, we need to extract some site covariates. This is so that we can see if there is systematic variation in the amount of intra-group clustering or repulsion depending on environmental factors at the site level.

If we think of this as a kind of linear model, we have our response of alpha, and intead of having site as a random variable now we include it as a dependent variable.

There are many factors that give difference between our sites as they span the continent of Australia. However, as all the sites are in tall eucalpt forest there will be some overall similarities in the environmental covariates. For forest productivity in Australia, we know that climate and soil is important. We can use the bioclim variables for climate as these cover what we want, we can use TWI, soil moisture, bulk density and sand for soil variables. We can also use data of the Forest Productivity Index (FPI), Gross Primary Productivity (GPP) and NDVI which takes into account temperature, solar radiation and precipitation to understand where productivity is highest. We can also calculate site-specific variables such as stand basal area and diversity indices.

### Make Site Covariate DF

```{r}
coords <- read.csv("../../data/ausplots/data_cleaned.csv")

site_covariates <- coords |> 
  select(Site_Name, Latitude, Longitude) |> 
  distinct()
```

### Site Coordinates

We need the coordinates of the sites first:

```{r}
coords <- coords |> 
  select(Latitude, Longitude) |> 
  distinct()
```

And then rename so we can run terra::extract

```{r}
coords <- coords |> 
  rename(lat = Latitude) |> 
  rename(lon = Longitude) |> 
  relocate(lon)
```

### Bioclim

Extract bioclim variables from geodata package

```{r}
clim <- worldclim_country(country = "Australia", 
                          var = "bio",
                          res = 0.5, 
                          path = tempdir())
```

Extract values for the sites and bind back into dataframe

```{r}
clim_vals <- extract(clim, coords)
```

And rename so column names are meaningful

```{r}
colnames(clim_vals) <- c("Site_Nam", 
                         "annual_mean_temp", 
                         "mean_dinural_range", 
                         "isothermality", 
                         "temp_seasonality", 
                         "max_temp_warmest_month", 
                         "min_temp_coldest_month", 
                         "temp_annual_range", 
                         "mean_temp_wettest_quart", 
                         "mean_temp_driest_quart", 
                         "mean_temp_warmest_quart", 
                         "mean_temp_coldest_quart", 
                         "annual_precip", 
                         "precip_wettest_month", 
                         "precip_driest_month", 
                         "precip_seasonality",
                         "precip_wettest_quart", 
                         "precip_driest_quart", 
                         "precip_warmest_quart",
                         "precip_coldest_quart")
```

Attach clim_vals back to site_covarites dataframe, preserving order:

```{r}
site_covariates <- cbind(site_covariates, clim_vals) 

site_covariates <- site_covariates[, -4] #remove ID col 
```

### Soil and Landscape

We can use the SLGA (the Soil and Landscape Grid of Australia) to pull out some soil and landscape variables that might be important. Keeping in mind that all sites have the **same** forest type (tall eucalypt forest) therefore there will be less variability in the types of soils and landscapes the sites are on.

Scraping SLGA 13/07/2026. Set the API key first to access COGs from url:

```{r}
Sys.setenv(
  "GDAL_HTTP_HEADER_FILE" = "tern_api_key.txt")
```

```{r}

slga_urls <- c('/vsicurl/https://data.tern.org.au/model-derived/slga/NationalMaps/SoilAndLandscapeGrid/DES/v2/DES_000_200_EV_N_P_AU_TRN_C_20190901.tif', #depth of soil (m)
               '/vsicurl/https://data.tern.org.au/model-derived/slga/NationalMaps/SoilAndLandscapeGrid/BDW/v2/BDW_000_005_EV_N_P_AU_TRN_N_20230607.tif', #bulk density (g/cm3) at 0-0.05m
               '/vsicurl/https://data.tern.org.au/model-derived/slga/NationalMaps/SoilAndLandscapeGrid/SND/v2/SND_000_005_EV_N_P_AU_TRN_N_20210902.tif', #sand % at 0-0.05m
               '/vsicurl/https://data.tern.org.au/model-derived/slga/NationalMaps/Covariates/Mosaics/90m/Clim_etaaann.tif', #average areal actual evapotranspiration annual
              '/vsicurl/https://data.tern.org.au/model-derived/slga/NationalMaps/Covariates/Mosaics/90m/Relief_twi_3s.tif') #twi

slga_stack <- rast(slga_urls)

#rename
names(slga_stack) <- c("depth_of_soil_m",
                     "bulk_density", 
                     "sand_pc", 
                     "annual_actual_ET",
                     "twi")

```

\

```{r}
#some urls aren't working properly, so download manually from TERN and add to the raster stack: 
#annual potential evaporation (mm) 
#organic carbon 
#total nitrogen % 0-0.05m 
#total phosphorous % 0-0.05m 

manual_stack <- rast(c("../../data/manual_tern_tifs/NTO_005_015_EV_N_P_AU_NAT_C_20231101.tif",
            "../../data/manual_tern_tifs/PTO_000_005_EV_N_P_AU_NAT_C_20231101.tif",
            "../../data/manual_tern_tifs/SOC_000_005_05_N_P_AU_TRN_N_20220727.tif", 
            "../../data/manual_tern_tifs/Clim_EPA.tif"))

#Add to other SLGA (TERN) data 
names(manual_stack) <- c("total_nitrogen", 
                         "total_phosphorous", 
                         "organic_carbon", 
                         "annual_PET")
#stack together and plot
slga_stack_full <- c(slga_stack, manual_stack)
plot(slga_stack_full, col = viridis(8, direction = -1))
```

And now we can extract the sites and bind back:

```{r}
slga_vals <- extract(slga_stack_full, coords)

site_covariates <- cbind(site_covariates, slga_vals)
site_covariates <- site_covariates[, -23] #remove ID col 
```

### Productivity

Forest productivity is hypothesised to affect the strength of competition between trees in forests. There are a range of productivity indices we can use as covariates.

The Forest Productivity Index (FPI), <https://data.gov.au/data/dataset/forest-productivity-index-version-1-0-sept-2024-release>, is one of these. It's based off ETA and PET climate data from BOM, NDVI, and 3PG modelling. The dataset is broken up into tiles of 37 across Australia and there is one tif for every year. So, what year? The censusing ended in 2015 we can choose this year. The different FPI rasters are different in extent, etc, so we cannot stack, but we can run a for loop instead:

```{r}
fpi_files <- c("../../data/forest_productivity_index/se55_2015_001.tif", 
           "../../data/forest_productivity_index/sh56_2015_001.tif",
           "../../data/forest_productivity_index/si50_2015_001.tif", 
           "../../data/forest_productivity_index/si56_2015_001.tif", 
           "../../data/forest_productivity_index/sj54_2015_001.tif",                               
           "../../data/forest_productivity_index/sj55_2015_001.tif",
           "../../data/forest_productivity_index/sk55_2015_001.tif")

working <- site_covariates[, 1:3]

for(i in fpi_files){
  r <- rast(i)
  
  r_vals <- extract(r, coords)
  r_vals <- r_vals[-1]
  working <- cbind(working, r_vals)
  
}


#and then summarise across columns to get to one column with all values 
working <- working |> 
  rowwise() |> 
  mutate(fpi = mean(c_across(4:10), 
                    na.rm = TRUE)) |> 
  ungroup()

#and attach this column to site_covariates df 

site_covariates <- cbind(site_covariates, working[11])
```

MODIS has data for Gross Primary Productivity (GPP).

Using the package MODISTools, we need to make a df in certain format, check what products and bands are available and then run the subset for our sites. Citation: Running, S., & Zhao, M. (2021). MODIS/Terra Net Primary Production Gap-Filled Yearly L4 Global 500m SIN Grid V061 \[Dataset\]. NASA Land Processes Distributed Active Archive Center <https://doi.org/10.5067/MODIS/MOD17A3HGF.061>. Date accessed: 2026-07-13

```{r}
df <- site_covariates[1:3] |> 
  rename(site_name = Site_Name) |> 
  rename(lat = Latitude) |> 
  rename(lon = Longitude)

mt_products()

mt_bands(product = "MOD17A3HGF") #yearly primary productivity
```

```{r}

pp <- mt_batch_subset(df = df, 
                      product = "MOD17A3HGF", 
                      band = "Npp_500m", 
                      start = "2015-01-01", 
                      end = "2015-12-", #extract from 2015 only
                      km_lr = 0, #just one point per site 
                      km_ab = 0)

```

\
And bind back to site covariates df:

```{r}
pp <- pp |> 
  select(site, value) |> 
  rename(Site_Name = site)

site_covariates <- left_join(site_covariates, 
                             pp, 
                             by = "Site_Name")

site_covariates <- site_covariates |> 
  rename(NPP = value)
```

### Diversity and Stand Basal Area

Inter-site variables can also be calculated. First, calculating the number of species in a site:

```{r}
data <- read.csv("../../data/ausplots/data_cleaned.csv")

num <- data |> 
  group_by(Site_Name) |> 
  summarise(c = n_distinct(Genus_Species)) |> 
  ungroup()
```

Calculating stand basal as DBH is given for all sites:

```{r}
data1 <- data |>  
  mutate(basal_area = (pi * ((Diameter/2)^2))/10000) |> 
  group_by(Site_Name) |> 
  summarise(sum(basal_area, na.rm = TRUE)) |> 
  ungroup()
```

And add these into larger dataframe:

```{r}
site_covariates <- left_join(site_covariates, num, by = "Site_Name")
site_covariates <- site_covariates |> 
  rename(no_species = c)
```

```{r}
site_covariates <- left_join(site_covariates, data1, by = "Site_Name")
site_covariates <- site_covariates |> 
  rename(basal_area_m2_ha = `sum(basal_area, na.rm = TRUE)`)
```

### Save site_covariates

First add the regions:

```{r}
site_covars <- site_covars |> 
  mutate(region = case_when(Site_Name %in% c("Carey", "Dombakup", "Warren", "Dawson", "Giants", "Sutton", "Frankland", "Clare", "Collins") ~ "WA", 
                            Site_Name %in% c("ANU363", "ANU589", "ANU101", "Ada Tree", "HardyCreek") ~ "VIC",
                            Site_Name %in% c("Turtons", "Lardner", "Weeaproinah") ~ "Otways", 
                            Site_Name %in% c("Newline", "WaratahMix", "WogWay", "Goodenia", "Candelo") ~ "seNSW", 
                            Site_Name %in% c("Baldy", "Herberton", "Lamb Range", "Koombooloomba") ~ "QLD", 
                            Site_Name %in% c("MinesRd", "A-Tree", "Tinebank", "Lorne", "BirdTree", "BlackBull", "Bruxner", "Osullivans") ~ "nNSW", 
                            Site_Name %in% c("Flowerdale", "Dip", "Mackenzie", "Caveside") ~ "ns_TAS", 
                            Site_Name %in% c("Bird", "Supersite", "NorthStyx", "Weld", "ZigZag", "MtField") ~ "ss_TAS", 
                            Site_Name %in% c("BlackRiver", "BondTier") ~ "King_TAS", 
                            Site_Name %in% c("MtMaurice", "BenRidge") ~ "Ben_TAS", 
                            TRUE ~ Site_Name))

site_covars <- site_covars |> 
  mutate(georegion = case_when(Site_Name %in% c("Carey", "Dombakup", "Warren", "Dawson", "Giants", "Sutton", "Frankland", "Clare", "Collins") ~ "WA", 
                               Site_Name %in% c("ANU363", "ANU589", "ANU101", "Ada Tree", "HardyCreek", "Turtons", "Lardner", "Weeaproinah") ~ "VIC", 
                               Site_Name %in% c("Newline", "WaratahMix", "WogWay", "Goodenia", "Candelo") ~ "seNSW", 
                               Site_Name %in% c("Baldy", "Herberton", "Lamb Range", "Koombooloomba") ~ "QLD", 
                               Site_Name %in% c("MinesRd", "A-Tree", "Tinebank", "Lorne", "BirdTree", "BlackBull", "Bruxner", "Osullivans") ~ "nNSW", 
                               Site_Name %in% c("Flowerdale", "Dip", "Mackenzie", "Caveside", "Bird", "Supersite", "NorthStyx", "Weld", "ZigZag", "MtField",       "BlackRiver", "BondTier", "MtMaurice", "BenRidge") ~ "TAS", 
                               TRUE ~ Site_Name))
```

And then save the site_covariates dataframe so we can explore in explore_site_covariates.qmd:

```{r}
write.csv(site_covariates, "site_covariates_new.csv")
```

### Check against old covariates

Previously I extracted the MAT and MAP from the ausplots supplementary materials. It would be good to check how similar the two data from the different sources are.

MAT is very similar, however MAP is changing quite a lot...

```{r}
new <- read.csv("scripts/analysis/site_covariates_new.csv")
old <- read.csv("data/site_covariates.csv")

old <- old |> select(Site_Name, MAT, MAP, FPI, GPP)
new <- new |> select(Site_Name, annual_mean_temp, annual_precip, fpi, NPP)

comb <- full_join(old, new, by = "Site_Name")

comb$MAP - comb$annual_precip

comb$MAT - comb$annual_mean_temp

```
