---
title: "Crown Class - Ausplot"
format: html
editor: visual
---

```{r}
library(dplyr)
library(ppjsdm)
library(ggbeeswarm)
library(spatstat.explore)
library(stringr)
```

## Ausplot - Crown Classes

Grouping individuals as eucalypt v non-eucalypt and then into their crown classes and then running ppjsdm analysisLoad Data

```{r}
data_cleaned <- read.csv("data/data_cleaned.csv")
```

### Source useful functions

```{r}
source(code/make_model_fun.R) 
source(code/make_summary_fun.R)
```

### Run models

Create vector of Site Names

```{r}
 sites <- unique(data_cleaned$Site_Name) 
 sites <- sites[!sites %in% c("Dip", "Bird", "Flowerdale")]

```

Run the crown class level model for all sites

```{r}
#make empty df to send results
full_df_com <- data.frame(matrix(ncol = 9, nrow = 0))
colnames(full_df_com) <- c("from", "to", "alpha", "lo", "hi", "lo_numerical",
                         "hi_numerical", "Potential", "site")

for(i in sites){ 
  
 ccmodel <- site_ccmodel(site = i, 
                         group_cd = TRUE, #should we group c and d together
                         threshold_noneuc = 8, #threshold to exclude non-euc groups
                         threshold_euc = 8, #threshold to exclude euc groups
                         short_range = 8) #run site_ccmodel function
 
 working_df <- make_sum_df(fits = list(ccmodel$fit), #use make_summary_df function to output summary as df 
                           summ = list(ccmodel$sum))
 
 working_df <- working_df %>% mutate(site = i) #set new column for site name 
 
 full_df_com <- rbind(full_df_com, working_df) #bind into larger df 
  
  }
  
```

I want to add how many individuals are in each group to the dataframe

```{r}
inds_full <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(inds_full) <- c("new_cc", "n", "site_name")

for(i in sites){ #do all sites at once and send into df 
  
  inds <- find_ind(site = i, 
                   group_cd = TRUE, 
                   threshold_noneuc = 8, 
                   threshold_euc = 8) #function above
  
  inds <- inds %>% mutate(site_name = i) #set new column called site_name for site 
  
  inds_full <- rbind(inds_full, inds) #bind into larger df 
}
```

Creating some new columns for easier plotting

```{r}
full_df_com <- full_df_com %>% 
  mutate(class_from = gsub("^[^ ]+ ", "\\1", from)) %>% 
  mutate(class_to = gsub("^[^ ]+ ", "\\1", to)) %>% #same for euc or non-euc 
  mutate(species_from = gsub("^([^ ]+).*", "\\1", from)) %>% 
  mutate(species_to = gsub("^([^ ]+).*", "\\1", to))

full_df_com <- full_df_com %>% ## and move site to the front to make it easier 
  relocate(site)
```

Rename inds_full dataframe to join it

```{r}
inds_full <- inds_full %>% 
  rename(to = from) 

full_df_com <- full_df_com %>%
  left_join(inds_full, by = c("site", "from")) 

full_df_com <- full_df_com %>% 
  rename(ind_from = n)
```

Lovely. I want to make sure that no unreasonably large confidence intervals exist in the dataframe. Unreasonably large are those CI's with a range more than 5 (this is as a strong effect is more than absolute 0.5).

```{r}
full_df_com <- full_df_com %>% 
  mutate(range_ci = hi - lo)
```

```{r}
large_ci <- full_df_com %>% 
  filter(range_ci > 5)
```

There are 2 problem sites: BirdTree and Collins. We need to go through each of them and rerun the model with the adjusted groups of individuals.

BirdTree

```{r}
birdtree <- data %>% filter(Site_Name == "BirdTree")

birdtree <- birdtree %>% 
  mutate(species = if_else(str_starts(Genus_Species, regex("^Eucalyptus|^Corymbia"), negate = TRUE), "Non-euc", "Eucalyptus")) %>% 
  mutate(new_cc = paste(species, Crown_Class))%>% 
  mutate(new_cc = if_else(new_cc == "Eucalyptus Emergent", "Eucalyptus Dominant", new_cc)) %>% 
  mutate(new_cc = if_else(new_cc %in% c("Eucalyptus Co-dominant", "Eucalyptus Dominant"), "Eucalyptus Co/dominant", new_cc)) %>% 
  mutate(new_cc = if_else(new_cc %in% c("Non-euc Co-dominant", "Non-euc Dominant"), "Non-euc Co/dominant", new_cc)) %>% 
  filter(!new_cc %in% c("Non-euc Co/dominant", "Non-euc Intermediate", "Eucalyptus Intermediate")) #Euc intermediate is the problem that wasn't removed last time 


birdtree %>%  count(new_cc)


birdtree <- birdtree %>%
  group_by(Ausplot_X, Ausplot_Y) %>% #group by coordinate columns
  mutate(
    is_duplicated = n() > 1, #create column of TRUE/FALSE 
    #new_column_name = if_else(condition, true, false): so condition=column name, if true=fill with, if false=fill with
    x_jitter = if_else(is_duplicated, Ausplot_X + runif(n(), -0.025, 0.025), Ausplot_X), #create x_jitter column
    y_jitter = if_else(is_duplicated, Ausplot_Y + runif(n(), -0.025, 0.025), Ausplot_Y) #create y_jitter column
  ) %>%
  ungroup() 


configuration<- Configuration(birdtree$x_jitter, birdtree$y_jitter, types = birdtree$new_cc) #make config 

# Set parameters
window <- ppjsdm::Rectangle_window(c(0, 100), c(0,100))
nspecies <- length(levels(configuration$types))
short_range <- matrix(8, nspecies, nspecies)
model <- "exponential"

fit_birdtree <- ppjsdm::gibbsm(configuration = configuration, #do the fit 
                           window = window,
                           short_range = short_range, 
                           model = "exponential", 
                           saturation = 10, 
                           nthreads = 4, 
                           fitting_package = "glmnet",
                           dummy_distribution = "stratified",
                           min_dummy = 1, dummy_factor = 1e10, 
                           max_dummy = 1e3)

sum_birdtree <- summary(fit_birdtree)
fit_birdtree$aic
```

Collins

```{r}
collins <- data %>% filter(Site_Name == "Collins")

collins <- collins %>% 
  mutate(species = if_else(str_starts(Genus_Species, regex("^Eucalyptus|^Corymbia"), negate = TRUE), "Non-euc", "Eucalyptus")) %>% 
  mutate(new_cc = paste(species, Crown_Class))%>% 
  mutate(new_cc = if_else(new_cc == "Eucalyptus Emergent", "Eucalyptus Dominant", new_cc)) %>% 
  mutate(new_cc = if_else(new_cc %in% c("Eucalyptus Co-dominant", "Eucalyptus Dominant"), "Eucalyptus Co/dominant", new_cc)) %>% 
  mutate(new_cc = if_else(new_cc %in% c("Non-euc Co-dominant", "Non-euc Dominant"), "Non-euc Co/dominant", new_cc)) %>% 
  filter(!new_cc ==  "Eucalyptus Intermediate") #Euc intermediate is the problem that wasn't removed last time 


collins %>%  count(new_cc)

configuration<- Configuration(collins$Ausplot_X, collins$Ausplot_Y, types = collins$new_cc) #make config 

# Set parameters
window <- ppjsdm::Rectangle_window(c(0, 100), c(0,100))
nspecies <- length(levels(configuration$types))
short_range <- matrix(8, nspecies, nspecies)
model <- "exponential"

fit_collins <- ppjsdm::gibbsm(configuration = configuration, #do the fit 
                           window = window,
                           short_range = short_range, 
                           model = "exponential", 
                           saturation = 10, 
                           nthreads = 4, 
                           fitting_package = "glmnet",
                           dummy_distribution = "stratified",
                           min_dummy = 1, dummy_factor = 1e10, 
                           max_dummy = 1e3)

sum_collins <- summary(fit_collins)
```

Put these three sites into a dataframe:

```{r}
df <- make_sum_df(fits = list(BirdTree = fit_birdtree, 
                              Collins = fit_collins), 
                  coefficient = "alpha",
                  summ = list(sum_birdtree, sum_collins))

#Have to make it the same as the full_df before rbind

df <- df %>% rename(site = Fit)

df <- df %>% 
  relocate(site) %>% 
  mutate(class_from = gsub("^[^ ]+ ", "\\1", from)) %>% 
  mutate(class_to = gsub("^[^ ]+ ", "\\1", to)) %>% #same for euc or non-euc 
  mutate(species_from = gsub("^([^ ]+).*", "\\1", from)) %>% 
  mutate(species_to = gsub("^([^ ]+).*", "\\1", to))


df <- df %>% 
  mutate(range_ci = hi - lo)

df %>% 
  filter(range_ci > 5) #check there are no large cis
```

Remove the three sites from the full_df

```{r}
full_df_com <- full_df_com %>% 
  filter(!site %in% c("BirdTree", "Collins"))

full_df_com %>% count(site)
```

Now make the final df

```{r}
final_ccmod <- rbind(full_df_com, df)

#And add a significance column
final_ccmod <- final_ccmod %>% mutate(sig = ifelse((lo > 0 | hi < 0), 1, NA ))
```

Save out df

```{r}
write.csv(final_ccmod, "final_ccmod.csv")
```

Make a euc interaction only df

```{r}
eucs <- final_ccmod %>% 
  filter(species_from == "Eucalyptus") %>% 
  filter(species_to == "Eucalyptus") %>% 
  mutate(class_int = paste0(class_from, sep ="_", class_to))
```

Need to add column for visualisation:\

```{r}
eucs <- eucs %>% 
  mutate(fill_col = ifelse(sig == 1, as.character(dom_euc), NA))
```

### Visualisation

Now we have a dataframe with all the data from the fits we can start visualisation

#### Scatterplots

```{r}
colours <- hcl.colors(7, "Viridis")

scatterplot <- ggplot() +
geom_vline(xintercept = 0, colour = "red", linewidth = .5) +
  geom_beeswarm(data = eucs, 
                aes(x = alpha, 
                    y = class_int, 
                    fill = fill_col, 
                    colour = dom_euc, 
                    text = paste("alpha:", alpha, "<br>", 
                                 "class int:", class_int, "<br>", 
                                 "dom euc:", dom_euc)), 
                 shape = 21, 
                size = 4,
                 alpha = 0.8) +
                  scale_color_manual(values = colours,
                                     name = "Dom Euc") + 
                  scale_fill_manual(values = colours,
                    na.value = "white", 
                    guide = "none") + 
  theme_bw() + 
   guides(colour = guide_legend(override.aes = 
                                 list(fill = colours, 
                                      size = 3) ) ) + 
  ylab("") 
scatterplot
```

```{r}
library(plotly)
ggplotly(scatterplot
         , tooltip = "text")
```

#### Box Plot

Suppressed - Suppressed

```{r}
#| fig-height: 6
sup <- eucs %>% 
  filter(class_int == "Suppressed_Suppressed") %>% 
  mutate(region = factor(region, levels = c("Ben_TAS", "King_TAS", "N_TAS", "S_TAS", "SE_VIC", "SE_NSW", "N_NSW", "QLD", "WA"))) %>%
mutate(site = fct_reorder(site, as.numeric(region)))

ggplot(sup, aes(x = alpha, 
                y = site, 
                colour = region, 
                fill = fill_col)) + 
  geom_vline(xintercept = 0, 
             colour = "red", 
             linetype = "dashed") +
  geom_errorbar(aes(xmin = lo, 
                    xmax = hi),
                size = 1.5) + 
  geom_point(shape = 21,
             size = 3) + 
  scale_colour_manual(values = c("#ED90A4", "#DC9E70", "#B7AE50", "#7EBA68", "#1DC199", "#00BFC8", "#6FB1E7", "#BB9DEA", "#E48FD1"), 
                      name = "Region") + 
  scale_fill_manual(values = c("Ben_TAS" = "#ED90A4","N_TAS"="#B7AE50","S_TAS" = "#7EBA68", "SE_NSW"= "#00BFC8","N_NSW" = "#6FB1E7","WA"=  "#E48FD1"), 
                    na.value = "white", 
                    guide = "none") +
  labs(x = "Alpha Coefficient (+/-95% CI)",
       y = "") + 
  scale_y_discrete(limits = rev) + #more making order right
  ggtitle("Suppressed_Suppressed") +
  theme_bw()  
  # guides(colour = guide_legend(override.aes = 
  #                                list(fill = colours, 
  #                                     size = 2) ) ) 
```

\

Co/dominant - Co/dominant

```{r}
colours <- hcl.colors(7, "Set2")

codom <- eucs %>% 
  filter(class_int == "Co/dominant_Co/dominant") %>% 
  mutate(fill_col = ifelse(sig == 1, as.character(dom_euc), NA)) %>% 
  arrange(dom_euc, alpha) %>% 
  mutate(site = factor(site, levels = unique(site)))


codom2 <- ggplot(codom, aes(x = alpha, y = site)) + 
    geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  geom_errorbar(aes(xmin = lo, xmax = hi, colour = dom_euc), size = 1.5) + 
  geom_point(aes(fill = fill_col, 
                 colour = dom_euc), 
             shape = 21,
             size = 3) +
  scale_colour_manual(values = c("#ED90A4", "#D3A263", "#99B657", "#33C192", "#00BDCE", "#94A9EC", "#DC91DB"), 
                      name = "Dominant Eucalyptus") + 
  scale_fill_manual(values = c("#ED90A4", "#99B657", "#33C192", "#00BDCE",  "#DC91DB"), 
                    na.value = "white", 
                    guide = "none")+
  labs(x = "Alpha Coefficient (+/-95% CI)",
       y = "") + 
  scale_y_discrete(limits = rev) + #more making order right
  ggtitle("Co/dominant_Co/dominant") +
  theme_bw() + 
  guides(colour = guide_legend(override.aes = 
                                 list(fill = colours, 
                                      size = 2) ) ) 

codom2
```

```{r}
codom <- eucs %>% 
  filter(class_int == "Co/dominant_Co/dominant") %>% 
  mutate(region = factor(region, levels = c("Ben_TAS", "King_TAS", "N_TAS", "S_TAS", "SE_VIC", "SE_NSW", "N_NSW", "QLD", "WA"))) %>%
mutate(site = fct_reorder(site, as.numeric(region)))

codoms <- ggplot(codom, aes(x = alpha, 
                y = site, 
                colour = region, 
                fill = fill_col)) + 
  geom_vline(xintercept = 0, 
             colour = "red", 
             linetype = "dashed") +
  geom_errorbar(aes(xmin = lo, 
                    xmax = hi),
                size = 1.5) + 
  geom_point(shape = 21,
             size = 3) + 
  scale_colour_manual(values = c("#ED90A4", "#DC9E70", "#B7AE50", "#7EBA68", "#1DC199", "#00BFC8", "#6FB1E7", "#BB9DEA", "#E48FD1"), 
                      name = "Region") + 
  scale_fill_manual(values = c("Ben_TAS" = "#ED90A4","King_TAS"= "#DC9E70", "S_TAS"= "#7EBA68", "SE_VIC"="#1DC199", "SE_NSW" = "#00BFC8", "N_NSW"= "#6FB1E7","QLD" =  "#BB9DEA", "WA"= "#E48FD1"), 
                    na.value = "white", 
                    guide = "none") +
  labs(x = "Alpha Coefficient (+/-95% CI)",
       y = "") + 
  scale_y_discrete(limits = rev) + #more making order right
  ggtitle("Co/dominant_Co/dominant") +
  theme_bw()  
codoms
```

Check what's happening between eucs and non-eucs

```{r}
non_euc <- final_ccmod %>% 
  mutate(species_int = paste0(species_from, sep = "_", species_to)) %>% 
  mutate(species_int = ifelse(species_int == "Eucalyptus_Non-euc", "Non-euc_Eucalyptus", species_int)) %>% 
  mutate(class_int = paste0(class_from, sep = "_", class_to)) %>% 
  mutate(sig = ifelse((lo > 0 | hi < 0), 1, NA )) %>% 
  filter(species_int == "Non-euc_Eucalyptus")
```

```{r}
colours <- hcl.colors(7, "Viridis")

ggplot() +
geom_vline(xintercept = 0, colour = "red", linewidth = .5) +
  geom_beeswarm(data = non_euc, 
                aes(x = alpha, 
                    y = class_int, 
                    fill = as.factor(sig), 
                    colour = dom_euc, 
                    text = paste("alpha:", alpha, "<br>", 
                                 "class int:", class_int, "<br>", 
                                 "dom euc:", dom_euc)), 
                 shape = 21, 
                 alpha = 0.8) +
                  scale_color_manual(values = colours,
                                     name = "Dom Euc") + 
                  scale_fill_manual(values = colours,
                    na.value = "white", 
                    name = "Sig", 
                    labels = c("1", "0")) + 
  theme_bw() + 
   guides(colour = guide_legend(override.aes = 
                                 list(fill = colours, 
                                      size = 2) ) ) + 
  ylab("") 

```

#### Prediction

Looking at prediction for some sites

```{r}
ppjsdm::plot_papangelou(fit_collins, 
                        window = window,
                type =  "Eucalyptus Suppressed", 
                use_log = TRUE, 
                legend_title = "Conditional Intensity",
                base_size = 16,
                drop_type_from_configuration = TRUE, 
                show =  "Eucalyptus Co/dominant")


```

### Model Performance

We need a systematic view of model performance across sites and crown classes. Therefore, we use conditional AUC to get the AUC score for each site for each group.

The function for conditional_auc:

Function for AUC for each site and group:

```{r}
#make empty df 
auc_df <- data.frame(matrix(ncol = 4))
colnames(auc_df) <- c("site", "types", "auc_score", "n_points")

auc_working <- auc_df
```

```{r}
auc_site <- function(site, 
                     threshold_noneuc = 12, #threshold to exclude non-euc groups
                     threshold_euc = 7, #threshold to exclude euc groups
                     short_range = 8){ #short range value 
  
d <- data_cleaned %>% filter(Site_Name == site)

d <- d %>% 
  mutate(species = if_else(str_starts(Genus_Species, regex("^Eucalyptus|^Corymbia"), negate = TRUE), "Non-euc", "Eucalyptus")) %>% 
  mutate(new_cc = paste(species, Crown_Class))%>% 
  mutate(new_cc = if_else(new_cc == "Eucalyptus Emergent", "Eucalyptus Dominant", new_cc)) %>% 
  mutate(new_cc = if_else(new_cc %in% c("Eucalyptus Co-dominant", "Eucalyptus Dominant"), "Eucalyptus Co/dominant", new_cc)) %>% 
  mutate(new_cc = if_else(new_cc %in% c("Non-euc Co-dominant", "Non-euc Dominant"), "Non-euc Co/dominant", new_cc))

d <- d %>%  #thresholds for groups 
  group_by(new_cc) %>% 
  mutate(observation_count = n()) %>% 
  filter(!(str_starts(new_cc, "Non-euc") & observation_count < threshold_noneuc)) %>% 
  filter(!(str_starts(new_cc, "Eucalyptus") & observation_count < threshold_euc)) 


counts <- d %>% count(new_cc)
#print(counts) # print counts to see 

d <- d %>%
  group_by(Ausplot_X, Ausplot_Y) %>% #group by coordinate columns
  mutate(
    is_duplicated = n() > 1, #create column of TRUE/FALSE 
    #new_column_name = if_else(condition, true, false): so condition=column name, if true=fill with, if false=fill with
    x_jitter = if_else(is_duplicated, Ausplot_X + runif(n(), -0.025, 0.025), Ausplot_X), #create x_jitter column
    y_jitter = if_else(is_duplicated, Ausplot_Y + runif(n(), -0.025, 0.025), Ausplot_Y) #create y_jitter column
  ) %>%
  ungroup() 


configuration<- Configuration(d$x_jitter, d$y_jitter, types = d$new_cc) #make config 

# Set parameters
window <- ppjsdm::Rectangle_window(c(0, 100), c(0,100))
nspecies <- length(levels(configuration$types))
short_range <- matrix(short_range, nspecies, nspecies)
model <- "exponential"

fit<- ppjsdm::gibbsm(configuration = configuration, #do the fit 
                           window = window,
                           short_range = short_range, 
                           model = "exponential", 
                           saturation = 10, 
                           nthreads = 4, 
                           fitting_package = "glmnet",
                           dummy_distribution = "stratified",
                           min_dummy = 1, dummy_factor = 1e10, 
                           max_dummy = 1e3)

model_types <- levels(configuration$types)

for(k in model_types){

  auc_working[k, "site"] <- site
  auc_working[k, "types"] <- k
   
p <- plot_papangelou(fit, #to find the AUC of the train data
                     window = window, 
                     drop_type_from_configuration = TRUE, 
                     type = k, 
                     use_log = FALSE, 
                     show = k, 
                     return_papangelou = TRUE)

  p$v[is.na(p$v)] <- mean(p)
  p$v[p$v == -Inf] <- -1e10

      
      X <- subset(as.ppp(configuration, W = window), marks == k)
      if(X$n <= 0){
        auc_working[k,"auc_score"] <- NA
      } else {
        auc_working[k, "auc_score"] <- auc(X = X, covariate = as.function(p))
      }
      
  auc_working[k, "n_points"] <- X$n
}

  return(auc_working)

}

```

```{r}
#test function
df <- auc_site(site = "BirdTree")

```

```{r}
#the sites 
sites <- unique(data_cleaned$Site_Name)
sites <- sites[!sites %in% c("Dip", "Bird", "Flowerdale")]
```

```{r}
for(i in sites){ 
  
  auc_working <- auc_site(site = i, 
                          threshold_noneuc = 8, 
                          threshold_euc = 8, 
                          short_range = 8)
  
auc_df <- rbind(auc_df, auc_working) #bind into larger df 
  
  }

```

Have a df, now to clean:

```{r}
auc_df_final <- data.frame(auc_df, row.names = NULL)
auc_df_final <- auc_df_final %>% 
  filter(!is.na(site))

auc_df_final %>% count(site)
```

Visualisation of spread of AUC scores

```{r}
#only care about euc scores so 
auc_df_final <- auc_df_final %>% 
  filter(str_starts(types, "E"))

auc_df_final %>% count(types)
         
```

```{r}
cc <- ggplot(data = auc_df_final, 
       aes(x = types, 
           y = auc_score)) + 
  geom_boxplot() + 
  geom_point(aes(colour = dom_euc)) + 
  ylim(0.4, 0.9) + 
  ggtitle("CC")
```

Summary statistics

```{r}
d <- final_ccmod %>% 
  filter(species_from == "Eucalyptus") %>% 
  filter(species_to == "Eucalyptus")

d <- d %>% 
  mutate(region = site) %>% 
  mutate(region = if_else(region %in% c("Carey", "Dombakup", "Warren", "Dawson", "Giants", "Sutton", "Frankland", "Clare", "Collins"), 
                          "WA", region)) %>% 
  mutate(region = if_else(region %in% c("ANU101", "ANU363", "ANU589", "Ada Tree", "HardyCreek", "Weeaproinah", "Turtons", "Lardner"), 
                          "SE_VIC", region)) %>% 
  mutate(region = if_else(region %in% c("Newline", "WaratahMix", "WogWay", "Goodenia", "Candelo"), 
                          "SE_NSW", region)) %>% 
  mutate(region = if_else(region %in% c("MinesRd", "A-Tree", "BirdTree", "BlackBull", "Lorne", "Tinebank", "Bruxner", "Osullivans"), 
                          "N_NSW", region)) %>% 
  mutate(region = if_else(region %in% c("Baldy", "Koombooloomba", "Lamb Range", "Herberton"), 
                          "QLD", region)) %>% 
  mutate(region = if_else(region %in% c("BenRidge", "Caveside", "Mackenzie", "MtField", "MtMaurice", "NorthStyx"), 
                          "d_TAS", region)) %>% 
  mutate(region = if_else(region %in% c("BondTier", "BlackRiver", "Weld", "Weld", "MtField", "ZigZag", "Supersite"), 
                          "o_TAS", region)) 


d <- d %>% filter(class_to == "Suppressed") %>% 
  filter(class_from == "Suppressed")

d %>% 
  group_by(region) %>% 
  summarise(
    count = n(), 
    median_a = median(alpha), 
    mean_a = mean(alpha), 
    min_a = min(alpha), 
    max_a = max(alpha)
  )

```

```{r}
### Model stats 
d <- final_ccmod %>% 
  filter(!species_from == "Non-euc") %>% 
  filter(!species_to == "Non-euc")

a <- d %>% count(sig)
d %>% count(region)

strong <- final_ccmod %>% 
  filter(alpha > 0.49 | alpha < -0.49)

```

```{r}
median(auc_df_final$auc_score)

auc_df_final %>% count(types)

d <- auc_df_final %>% 
  filter(types %in% c("Eucalyptus Co/dominant", "Eucalyptus Intermediate", "Eucalyptus Suppressed"))

median(d$auc_score)
```

```{r}
########### Cross validation
#As we know, the AUC method used above doesn't allow for any within-group interactions to be evaluated 
#Therefore we can use leave-n-individuals out cross validation

crownclass <- function(site, 
                      threshold = 10){ #short range value 
 d <- data_cleaned %>% filter(Site_Name == site)

d <- d %>% 
  mutate(species = if_else(str_starts(Genus_Species, regex("^Eucalyptus|^Corymbia"), negate = TRUE), "Non-euc", "Eucalyptus")) %>% 
  mutate(new_cc = paste(species, Crown_Class))%>% 
  mutate(new_cc = if_else(new_cc == "Eucalyptus Emergent", "Eucalyptus Dominant", new_cc)) %>% 
  mutate(new_cc = if_else(new_cc %in% c("Eucalyptus Co-dominant", "Eucalyptus Dominant"), "Eucalyptus Co/dominant", new_cc)) %>% 
  mutate(new_cc = if_else(new_cc %in% c("Non-euc Co-dominant", "Non-euc Dominant"), "Non-euc Co/dominant", new_cc))

d <- d %>%  #thresholds for groups 
  group_by(new_cc) %>% 
  mutate(observation_count = n()) %>% 
  filter(!(str_starts(new_cc, "Non-euc") & observation_count < threshold)) %>% 
  filter(!(str_starts(new_cc, "Eucalyptus") & observation_count < threshold)) 


counts <- d %>% count(new_cc)
#print(counts) # print counts to see 

d <- d %>%
  group_by(Ausplot_X, Ausplot_Y) %>% #group by coordinate columns
  mutate(
    is_duplicated = n() > 1, #create column of TRUE/FALSE 
    #new_column_name = if_else(condition, true, false): so condition=column name, if true=fill with, if false=fill with
    x_jitter = if_else(is_duplicated, Ausplot_X + runif(n(), -0.025, 0.025), Ausplot_X), #create x_jitter column
    y_jitter = if_else(is_duplicated, Ausplot_Y + runif(n(), -0.025, 0.025), Ausplot_Y) #create y_jitter column
  ) %>%
  ungroup()

}
```

```{r}
ind_validation3 <- function(site, 
                            n_ind = 10, #number of individuals to leave out 
                            n_it = 30) #number of iterations, default to 30
{
  
  ### Make empty dataframes 
  auc_train_ind <- data.frame(matrix(ncol = 3, nrow = 0))
  auc_test_ind <- data.frame(matrix(ncol = 3, nrow =0))
  colnames(auc_train_ind) <- c("iter", "types", "value")
  colnames(auc_test_ind) <- c("iter", "types", "value")
  
  auc_train_ind_working <- auc_train_ind
  auc_test_ind_working <- auc_test_ind
  
  data <- crownclass(site,  #function to get the site data into new groups 
             threshold = 10)
  
  #get the types from the model
  types <- data %>% 
    pull(new_cc) %>% 
    unique() %>% 
    sort()
  
  ## If there is less than 15 individuals in a group, we do not run the AUC for that group
  model_types <- data %>% 
    group_by(new_cc) %>% 
    mutate(count = n()) %>% 
    ungroup() %>% 
    filter(count > 15) %>% ### !!! If there is less than 15 individuals in a group, the AUC is not done
    pull(new_cc) %>% 
    unique() %>% 
    sort()
  
  nspecies <- length(types)
  window <- ppjsdm::Rectangle_window(c(0, 100), c(0,100))
  
  #Training loop 
  for (i in 1:n_it) {
    
    set.seed(i)
    print(paste0("Starting n = ", i))
    
    
    for (j in model_types){
      
      
      #### Select 20 individuals from mark j for iteration i
      df_test <- data %>% filter(new_cc == j) 
      df_test <- df_test[sample(nrow(df_test), n_ind, replace = FALSE), ]
      df_train <- anti_join(data, df_test)
      # semi_join(df, leave_out_df) #check if the anti_join has worked 
      
      
      #### Make test and train configurations from sampled datasets 
      configuration_train <- Configuration(df_train$x_jitter, df_train$y_jitter, df_train$new_cc) 
      configuration_test <- Configuration(df_test$x_jitter, df_test$y_jitter, df_test$new_cc)
      
      ### Fit the training model
      fit_train <- ppjsdm::gibbsm(configuration_train, 
                                  window = window, 
                                  short_range = matrix(8, nspecies, nspecies), 
                                  model = "exponential",
                                  fitting_package = "glmnet",
                                  saturation = 10, 
                                  dummy_distribution = "stratified",
                                  min_dummy = 1, dummy_factor = 1e10, 
                                  max_dummy = 1e3) 
      
      print(paste0("*** Starting model = ", j))
      
      
      
      auc_train_ind_working[1, "iter"] <- i
      auc_train_ind_working[1, "types"] <- j
      
      p_train <- plot_papangelou(fit_train, #predict to training model fit 
                                 drop_type_from_configuration = TRUE, 
                                 type = j, 
                                 use_log = FALSE, 
                                 show = j, 
                                 return_papangelou = TRUE)
      
      p_train$v[is.na(p_train$v)] <- mean(p_train)
      p_train$v[p_train$v == -Inf] <- -1e10
      
      # AUC values for training fit 
      X <- subset(as.ppp(configuration_train, W = window), marks == j) 
      auc_train_ind_working[1, "value"] <- auc(X, covariate = as.function(p_train)) 
      
      auc_train_ind <- rbind(auc_train_ind, auc_train_ind_working)
      
      
      #Testing/Predicting 
      auc_test_ind_working[1, "iter"] <- i
      auc_test_ind_working[1, "types"] <- j
      
      #Using fit_train to predict to configuration_test
      p <- plot_papangelou(fit_train, 
                           configuration = configuration_train,
                           type = j,
                           drop_type_from_configuration = FALSE,
                           use_log = FALSE, 
                           return_papangelou = TRUE)
      
      X <- subset(as.ppp(configuration_test, W = window), marks == j)
      X <- ppp(x = X$x,
               y = X$y,
               window = X$window)
      auc_test_ind_working[1, "value"] <- auc(X = X, covariate = as.function(p)) #testing auc
      
      
      auc_test_ind <- rbind(auc_test_ind, auc_test_ind_working)
      
    } # closing the j loop
    
  }# closing i loop
  
  
  
  auc_train_ind$split <- "train"
  auc_test_ind$split <- "test"
  
  rbind(auc_train_ind, auc_test_ind)
  
} #closing function 

```

```{r}
## For all sites: 
sites <- unique(data_cleaned$Site_Name) 
sites <- sites[!sites %in% c("Dip", "Bird", "Flowerdale")]

     
                 
      
        
      
      

full_auc_ind <- data.frame(matrix(ncol = 5, nrow = 0))
colnames(full_auc_ind) <- c("iter", "types", "value", "split", "site")


a <- ind_validation3(site = "BenRidge", 
                     n_ind = 10, 
                     n_it = 20)
  
  a$site <- "BenRidge" #set column site to k 
  
  full_auc_ind <- rbind(full_auc_ind, a) #bind into larger df 
  
  full_auc_ind %>% count(site)
```

```{r}
write.csv(full_auc_ind, "auc_inds_cc.csv")

d <- full_auc_ind %>% 
  filter(!types %in% c("Non-euc Co/dominant", "Non-euc Intermediate", "Non-euc Suppressed"))

auc_inds_sum <- full_auc_ind %>% 
  group_by(types, split) %>% 
  summarise(
    median_a = median(value), 
    mean_a = mean(value), 
    min_a = min(value), 
    max_a = max(value)
  )
```
